### 2016-05-04 writing/latex/wilf.tex
spell occurence correctly, add oeis seqs

\section{Wilf-classes with patterns of length 1.}
When considering the mesh patterns of length \(2\) it will be useful to know
the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
\(\av{\perm{2,3,1}}\), this means that we are consdering the set \(\av{\{231,p\}}\)
where \(p\) is a mesh-pattern of length 1.

@@

\nextvar
The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
quadrant marked mesh patterns studied by Remmel,
Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
avoiders by decomposing a non-empty avoider of \(p_1\) around the maximum element
in order to give the following structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\). The lower-left section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

Evauluation of this generating function gives the Fine numbers
(\href{https://oeis.org/A000957}{OEIS: A000957}).
\begin{equation*}
    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
\end{equation*}

@@

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(r\) must fit into precisely on of the following twp forms.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, minimum cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function is enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364} }
\end{equation*}

@@

In order to consider symmetries we must only take the symmetries that preserve
the pattern \(\perm{2,3,1}\). The only symmetry that preserves the pattern
\(\perm{2,3,1}\) is \emph{reverse-complement-inverse}. Using this
symmetry to merge classes gives us \(61\) classes of trivial
Wilf-equivalences, these Wilf-equivalences are explained by patterns being either
coincident, or being the reverse-complement-inverse of a pattern.

@@

For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the region corresponding to the unshaded box in each case. Each
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
\end{equation*}

@@

\subsection{}
The containers of the following patterns
have exactly one occurrence.
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Once again we consider the regions corresponding to the unshaded regions, For each
pattern
except the first the two regions are independent, and one contains
any avoider of \(\perm{2,3,1}\) and the other must contain a
decreasing sequence by \LemmaRef{lem:incdec}.
Let us consider the first pattern separately. In order to
avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
we can add some additional restrictions, \ie all elements in the top region must
be to the right of all elements in the bottom region.
\begin{equation}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
        \draw (1.5,0) -- (1.5,1)
              (1.5,2) -- (1.5,3);
        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \end{tikzpicture}
    }
    \label{eqn:splitboxes}
\end{equation}
Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
two do not interact in any manner. The containers of this pattern are counted
the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
The containers
have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
therefore gives us
\begin{equation*}
    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
\end{equation*}

@@

Therefore these two patterns have been shown to have the same number of
avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
The avoiders have general form
\begin{equation*}
  \var_k = C_k - (k-1), \var_0 = 1
\end{equation*}
and have enumeration
\begin{equation*}
    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc
\end{equation*}

@@

Considering an occurrence of either of these patterns in a permutation  we know
the following about the points corresponding to the points in the patterns.
\begin{itemize}
    \item The point corresponding to the first point in both patterns
        must lie in the first block of the set partition (there are
        no points southwest from it in the permutation).
    \item The point corresponding to the second point in both patterns
        is a block bottom (there are no points southeast of it in
        the permutation).
    \item If the region corresponding to box \((2,2)\) in an occurrence of
    \(m_1\) is empty, then the point corresponding to the second point is
    precisely the last block bottom. If the region corresponding to box \((0,1)\)
    in an occurrence of \(m_2\) is empty, then the point corresponding to the
    second point is precisely the first block bottom.
    If these regions are non-empty then the block containing the point
    corresponding to the second point in both patterns contains only the point
    (it is a singleton block).
\end{itemize}
This tells us that an occurence of the patterns must happen when there
is a singleton block occurring after the first block. The difference
between the patterns is in the underlying classical pattern. This
means that permutations containing \(m_1\) correspond to set partitions
with a singleton block with value one higher than some element in the
block containing \(1\). The permutations containing \(m_2\) correspond
to the set partitions containing a block with block bottom having value
one lower than some element in the block containing \(1\) and if this
block is not the block containing \(1\) then it is a singleton block. This
proves that the containers of both of these patterns in \(\av{231}\) are
equinumerous, and therefore so are their avoiders.

@@

First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\).
We can perform the usual structural decomposition of an avoider
of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
% Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
% precisely once.
If \(\scriptvar_1\) is the set \(\av{\perm{2,3,1},m_1}\) then any permutation in
\(\scriptvar_1\) either starts with a maximum or does not, giving us the
decomposition
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}

Now consider the decomposition around the maximum of a permutation in
\(\scriptvar_2 = \av{\perm{2,3,1},m_2}\), the permutation either ends with
the maximum, or it does not, so we get
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
    \decompmax{0/1,1/0,1/1}{\mathcal{\varfine}}{}{}{}
\end{equation*}
Therefore both of these sets of avoiders are enumerated in the same manner
having generating function satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x\varfine(x)
\end{equation*}
This generating function gives
\begin{equation*}
    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc
\end{equation*}

@@

First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
around the maximum. We have different conditions if we start with the maximum or not.
\begin{equation*}
    \scriptvar_1 =
    \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
\end{equation*}

Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
conditions for starting with the maximum
\begin{equation*}
    \scriptvar_2 =
    \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
    \sqcup
    \decompleft{0/1,1/0,1/1}{\scriptvar_2}{}{}{}
\end{equation*}
This gives us two generating functions satisfying the following pair of equations
\begin{align}
    \var_1(x) &= 1 + xC(x)(\varmaxl(x)-1) +x\var_1(x) \label{eqn:genL1}\\
    \text{ and }\var_2(x) &= 1 + x(C(x)-1)^2 + x\var_2(x) \label{eqn:genL2}
\end{align}
In order for these two functions to give the same value it is necessary to show that
\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
\(C(x)(\varmaxl(x)-1) = (C(x)-1)^2\).
\begin{align*}
    &&(C(x)-1)^2 &= C(x)(\varmaxl(x)-1)\\
        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(\varmaxl\) and \(C\)}\\
        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
\end{align*}
The final line is always satisfied since it is the form of \(C(x)\), and
therefore the two generating functions are equal.

@@

Evaluating the generating function \(\var(x)\) gives us the enumeration
\begin{equation*}
    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc
\end{equation*}

@@

The avoiders of the third pattern can be decomposed by the maximum
to give
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{\varfine}}{}{}{\scriptvar_3}
\end{equation*}
The generating function derived satisfies \(\var_3(x) = 1 + x\varfine(x)\var_3(x)\).
The fourth pattern can be decomposed around the maximum in a similar
manner.
\begin{equation*}
    \scriptvar_4 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptvar_4}{}{}{\mathcal{\varfine}}
\end{equation*}

We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
        &= \frac{1-xC(x) +x}{1-xC(x)}\\
        &= 1+xC(x)
    \end{aligned}
\end{equation*}
We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
 and have enumeration sequence
 \begin{equation*}
     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc
 \end{equation*}
 This is an offset of the Catalan numbers.

@@

Now we consider decomposition of an avoider
\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2}  in \(\av{231}\) around the maximum,
this avoider has form
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{\varmaxl}_2}{}{}{\mathcal{C}}
\end{equation*}
This gives us the generating function
\begin{equation*}
    \begin{aligned}
        \var_2(x) &= 1+xC(x)\varmaxl(x)\\
        &= 1+xC(x)(1+xC(x)(C(x)-1))\\
        &= 1 +xC(x)(C(x)-xC(x)) && C(x) = 1+xC^2(x)\\
        &= 1 +x + xC^2(x) - xC(x) \\
        &= 1 +x + xC(x)(C(x)-1)
    \end{aligned}
\end{equation*}
Therefore this generating function is the same as \EquationRef{eq:qgen}.
We can decompose
\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/1},
\textpattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1},
\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/0},
\textpattern{}{1,2}{0/0,0/1,0/2,1/2,2/0,2/1,2/2},
and \textpattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2} around the leftmost point
into an avoider of one of the patterns with generating function \(\varmaxl(x)\) and
an avoider of \(\perm{2,3,1}\).

Now decompose an avoider of \textpattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1}
around the leftmost point.
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_3\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup \decompleft{0/0,0/1,1/0}{}{}{}{\varmaxl_3}
\end{equation*}
This gives generating function \(\var_3(x)\) satisfying
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 +xC(x)(\var_3(x)-1) + x\varmaxl(x)\\
        &= C(x) - xC^2(x) + xC(x) +x^2C^3(x) - x^2C^2 && \text{Solving for \(\var_3(x)\)}\\
        &= 1 + xC(x) +x^2C^3(x)-xC^2(x) && C = 1+xC^2(x)\\
        &= 1 + x + x^2C^3(x) && xC^2(x) = C(x)-1\\
        &= 1 + x + xC(x)(C(x)-1)
    \end{aligned}
\end{equation*}
This is equivalent to \EquationRef{eq:qgen}, and therefore
the classes containing all of these patterns are Wilf-equivalent.
The classes have enumeration
\begin{equation*}
    1, 1, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
\end{equation*}
This is an offset of OEIS Sequence number \href{http://oeis.org/A071724}{A071724}.

>>>>

\section{Wilf-classes with patterns of length 1.}
When considering the mesh patterns of length \(2\) it will be useful to know
the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
\(\av{\perm{2,3,1}}\), this means that we are considering the set \(\av{\{231,p\}}\)
where \(p\) is a mesh-pattern of length 1.

@@

\nextvar
The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
quadrant marked mesh patterns studied by \textcite{2012arXiv1201.6243K}.
Alternatively we can enumerate
avoiders by decomposing a non-empty avoider of \(p_1\) around the maximum element
in order to give the following structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\). The lower-left section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

Evaluation of this generating function gives the Fine numbers
(\href{https://oeis.org/A000957}{OEIS: A000957}).
\begin{equation*}
    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
\end{equation*}

@@

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(r\) must fit into precisely on of the following two forms.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, minimum cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function is enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364}}
\end{equation*}

@@

In order to consider symmetries we must only take the symmetries that preserve
the pattern \(\perm{2,3,1}\). The only symmetry that preserves the pattern
\(\perm{2,3,1}\) is \emph{reverse-complement-inverse}. Using this
symmetry to reduces the number of Wilf-classes gives us \(61\) classes of trivial
Wilf-equivalences, these Wilf-equivalences are explained by patterns being either
coincident, or being the reverse-complement-inverse of a pattern.

@@

For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the region corresponding to the unshaded box in each case. Each
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n\) - \href{https://oeis.org/A001453}{A001453} offset 2}
\end{equation*}

@@

\subsection{}
The containers of the following patterns
have exactly one occurrence.
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Once again we consider the regions corresponding to the unshaded regions, For each
pattern
except the first the two regions are independent, and one contains
any avoider of \(\perm{2,3,1}\) and the other must contain a
decreasing sequence by \LemmaRef{lem:incdec}.
Let us consider the first pattern separately. In order to
avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
we can add some additional restrictions, \ie all elements in the top region must
be to the right of all elements in the bottom region.
\begin{equation}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
        \draw (1.5,0) -- (1.5,1)
              (1.5,2) -- (1.5,3);
        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \end{tikzpicture}
    }
    \label{eqn:splitboxes}
\end{equation}
Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
two do not interact in any manner. The containers of this pattern are counted
the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
The containers
have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
therefore gives us
\begin{equation*}
    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A014137}{A014137} offset 2}
\end{equation*}

@@

Therefore these two patterns have been shown to have the same number of
avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
The avoiders have general form
\begin{equation*}
  \var_k = C_k - (k-1), \var_0 = 1
\end{equation*}
and have enumeration
\begin{equation*}
    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A000027}{A000027} offset 2}
\end{equation*}

@@

Considering an occurrence of either of these patterns in a permutation  we know
the following about the points corresponding to the points in the patterns.
\begin{itemize}
    \item The point corresponding to the first point in both patterns
        must lie in the first block of the set partition (there are
        no points southwest from it in the permutation).
    \item The point corresponding to the second point in both patterns
        is a block bottom (there are no points southeast of it in
        the permutation).
    \item If the region corresponding to box \((2,2)\) in an occurrence of
    \(m_1\) is empty, then the point corresponding to the second point is
    precisely the last block bottom. If the region corresponding to box \((0,1)\)
    in an occurrence of \(m_2\) is empty, then the point corresponding to the
    second point is precisely the first block bottom.
    If these regions are non-empty then the block containing the point
    corresponding to the second point in both patterns contains only the point
    (it is a singleton block).
\end{itemize}
This tells us that an occurrence of the patterns must happen when there
is a singleton block occurring after the first block. The difference
between the patterns is in the underlying classical pattern. This
means that permutations containing \(m_1\) correspond to set partitions
with a singleton block with value one higher than some element in the
block containing \(1\). The permutations containing \(m_2\) correspond
to the set partitions containing a block with block bottom having value
one lower than some element in the block containing \(1\) and if this
block is not the block containing \(1\) then it is a singleton block. This
proves that the containers of both of these patterns in \(\av{231}\) are
equinumerous, and therefore so are their avoiders.

@@

First consider the structure of an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\).
We can perform the usual structural decomposition of an avoider
of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
% Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
% precisely once.
If \(\scriptvar_1\) is the set \(\av{\perm{2,3,1},m_1}\) then any permutation in
\(\scriptvar_1\) either starts with a maximum or does not, giving us the
decomposition
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}
Where \(\mathcal{\varfine} = \av{(231,\textpattern{}{1}{0/1,1/0})}\).
Now consider the decomposition around the maximum of a permutation in
\(\scriptvar_2 = \av{\perm{2,3,1},m_2}\), the permutation either ends with
the maximum, or it does not, so we get
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
    \decompmax{0/1,1/0,1/1}{\mathcal{\varfine}}{}{}{}
\end{equation*}
Therefore both of these sets of avoiders are enumerated in the same manner
having generating function satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x\varfine(x)
\end{equation*}
This generating function gives
\begin{equation*}
    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A000958}{A000958} offset 2}
\end{equation*}

@@

First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
around the maximum. We have different conditions if we start with the maximum or not.
\begin{equation*}
    \scriptvar_1 =
    \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
\end{equation*}
Where \(\mathcal{\varmaxl}_1 = \av{(231,\textpattern{}{1}{0/0,0/1,1/1})}\).
Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
conditions for starting with the maximum
\begin{equation*}
    \scriptvar_2 =
    \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
    \sqcup
    \decompleft{0/1,1/0,1/1}{\scriptvar_2}{}{}{}
\end{equation*}
This gives us two generating functions satisfying the following pair of equations
\begin{align}
    \var_1(x) &= 1 + xC(x)(\varmaxl(x)-1) +x\var_1(x) \label{eqn:genL1}\\
    \text{ and }\var_2(x) &= 1 + x(C(x)-1)^2 + x\var_2(x) \label{eqn:genL2}
\end{align}
In order for these two functions to give the same value it is necessary to show that
\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
\(C(x)(\varmaxl(x)-1) = (C(x)-1)^2\).
\begin{align*}
    &&(C(x)-1)^2 &= C(x)(\varmaxl(x)-1)\\
        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(\varmaxl\) and \(C\)}\\
        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
\end{align*}
The final line is always satisfied since it is the form of \(C(x)\), and
therefore the two generating functions are equal.

@@

Evaluating the generating function \(\var(x)\) gives us the enumeration
\begin{equation*}
    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc\tag{\(C_n - \) \href{https://oeis.org/A014138}{A014138} offset 1}
\end{equation*}

@@

The avoiders of the third pattern can be decomposed by the maximum
to give
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{\varfine}}{}{}{\scriptvar_3}
\end{equation*}
Where \(\mathcal{\varfine} = \av{(231,\textpattern{}{1}{0/1,1/0})}\).
The generating function derived satisfies \(\var_3(x) = 1 + x\varfine(x)\var_3(x)\).
The fourth pattern can be decomposed around the maximum in a similar
manner.
\begin{equation*}
    \scriptvar_4 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptvar_4}{}{}{\mathcal{\varfine}}
\end{equation*}

We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
        &= \frac{1-xC(x) +x}{1-xC(x)}\\
        &= 1+xC(x)
    \end{aligned}
\end{equation*}
We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
 and have enumeration sequence
 \begin{equation*}
     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc \tag{\href{https://oeis.org/A000108}{OEIS: A000108} offset 1}
 \end{equation*}
 This is an offset of the Catalan numbers.

@@

Now we consider decomposition of an avoider of
\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2}  in \(\av{231}\) around the maximum,
this avoider has form
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{\varmaxl}_2}{}{}{\mathcal{C}}
\end{equation*}
Where \(\mathcal{\varmaxl}_2 = \av{(231,\textpattern{}{1}{0/1,1/0,1/1})}\).
This gives us the generating function
\begin{equation*}
    \begin{aligned}
        \var_2(x) &= 1+xC(x)\varmaxl(x)\\
        &= 1+xC(x)(1+xC(x)(C(x)-1))\\
        &= 1 +xC(x)(C(x)-xC(x)) && C(x) = 1+xC^2(x)\\
        &= 1 +x + xC^2(x) - xC(x) \\
        &= 1 +x + xC(x)(C(x)-1)
    \end{aligned}
\end{equation*}
Therefore this generating function is the same as \EquationRef{eq:qgen}.
We can decompose
\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/1},
\textpattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1},
\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/0},
\textpattern{}{1,2}{0/0,0/1,0/2,1/2,2/0,2/1,2/2},
and \textpattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2} around the leftmost point
into an avoider of one of the patterns with generating function \(\varmaxl(x)\) and
an avoider of \(\perm{2,3,1}\).

Now decompose an avoider of \textpattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1}
around the leftmost point.
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_3\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup \decompleft{0/0,0/1,1/0}{}{}{}{\varmaxl_3}
\end{equation*}
Where \(\mathcal{\varmaxl}_3 = \av{(231,\textpattern{}{1}{0/0,0/1,1/0})}\).
This gives generating function \(\var_3(x)\) satisfying
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 +xC(x)(\var_3(x)-1) + x\varmaxl(x)\\
        &= C(x) - xC^2(x) + xC(x) +x^2C^3(x) - x^2C^2 && \text{Solving for \(\var_3(x)\)}\\
        &= 1 + xC(x) +x^2C^3(x)-xC^2(x) && C = 1+xC^2(x)\\
        &= 1 + x + x^2C^3(x) && xC^2(x) = C(x)-1\\
        &= 1 + x + xC(x)(C(x)-1)
    \end{aligned}
\end{equation*}
This is equivalent to \EquationRef{eq:qgen}, and therefore
the classes containing all of these patterns are Wilf-equivalent.
The classes have enumeration
\begin{equation*}
    1, 1, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A071724}{OEIS: A071724 with offset}}
\end{equation*}
